Sizing the bets in a focused portfolio

### 2024-02-23 doc/informal/paper.tex
Formal paper (for submission to arXiv)

The paper attempts to provide a mathematical model and a tool for the focused
investing strategy as advocated by Buffett~\cite{berkshireLetters},
Munger~\cite{almanack}, and others from this investment community.
The approach presented here assumes that the investor's role is to think about
probabilities of different outcomes for a set of businesses. Based on these
assumptions, the tool calculates the optimal allocation of capital for
each of the investment candidates. The model has the option to provide
constraints that ensure: no shorting, no use of leverage, providing a maximum
limit to the risk of permanent loss of capital, and maximum individual
allocation. The software is made available for public use.\\

This paper consist of two parts, one without the mathematics and one with the
mathematics. For a non--mathematical reader, it is best to skip the two sections
related to mathematics and numerics, namely~\autoref{sec:mathematics}
and~\autoref{sec:numerics}. The reader who is interested in mathematics should
still read the preceding sections since they provide context within which the
mathematics is built on. Moreover, for mathematically skilled reader, the
problem formulated and solved here represents a non--linear constrained
optimization problem.

@@

The answer to this question has been given by Kelly with his widely known Kelly
formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
starts by maximizing long--term growth of capital when one is presented with
an infinite amount of opportunities to bet on. This work extends that idea by
considering:
\begin{enumerate}
    \item Multiple companies (stocks) in parallel,
    \item An arbitrary number of scenarios for each company,
    \item Multiple constraints for modelling our preference towards no shorting,
    no use of leverage, and providing a maximum value for the risk of permanent
    loss of capital,
    \item A fundamentals--based analysis with a very long time horizon. After
    all, a stock is an ownership share of business~\cite{intelligentInvestor},
    in which time--scales are significantly larger than the time--scales of
    day--to--day stock price fluctuations.
\end{enumerate}

@@

During the mathematical derivation presented in \autoref{sec:mathematics}, an
assumption is made that the number of bets is very high (tends to infinity).
This work does not try to justify this assumption in a strong mathematical
form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
soft, non-mathematical reasoning on why the authors think this is fine: The
assumption is made in order to write the equations in terms of probabilities
instead of the number of outcomes divided by the number of bets. Therefore, as
long as the input probabilities are \textit{conservatively} estimated, the
framework should still be valid. This essentially represents the most important
margin of safety~\cite{intelligentInvestor}.

@@

    \item Maximum allowable permanent capital loss constraint ensures that the
    worst--case outcome does not exceed losing a specified amount of capital
    with a specified probability. Adding this constraint means: "Under a
    worst--case scenario, I am comfortable permanently losing $K$ (e.g. 25\%) of
    my capital with probability $P$ (e.g. 0.1\%).":
        \begin{equation}
        \label{eq:14}
            \sum_j^{N_c} f_j \min(p_{ij} k_{ij}) \ge P \cdot K \rightarrow
            -\sum_j^{N_c} \left(f_j \min(p_{ij} k_{ij}) \right) + P \cdot K + s = 0
        \end{equation}
    \noindent where $\min(p_{ij} k_{ij})$ is the worst--case outcome across all
    scenarios for the $j$-th candidate company. Here, the minimum returns
    $k_{ij}$ and the maximum worst--case return $K$ are both negative by
    convention, indicating a loss, hence the $\ge$ sign.

@@

\end{enumerate}

@@

\noindent The observation about excessive diversification is inline with the
thoughts from the Poor Charlie's Almanack~\cite{almanack} and one of the
lectures from Li Lu that the authors frequently watch~\cite{liLuLecture}.

>>>>

The paper attempts to provide a mathematical model and a tool for the focused
investing strategy as advocated by Buffett~\cite{berkshireLetters},
Munger~\cite{almanack}, and others from this investment community.
The approach presented here assumes that the investor's role is to think about
probabilities of different outcomes for a set of businesses. Based on these
assumptions, the tool calculates the optimal allocation of capital for
each of the investment candidates. The model is based on a generalized Kelly
Criterion with options to provide constraints that ensure: no shorting, no use
of leverage, providing a maximum limit to the risk of permanent loss of capital,
and maximum individual allocation. The software is applied to an example
portfolio from which certain observations about excessive diversification are
obtained. In addition, the software is made available for public use.\\

@@

The answer to this question has been given by Kelly with his widely known Kelly
Criterion~\cite{kelly}. Despite the fact that the Kelly Criterion has been
published in 1956 in the information theory, its use is now widespread
in the field of finance and sports betting, with a plethora of recent research.
For example, Pinelis and Ruppret~\cite{pinelisRuppert} noted that the
asset allocation fractions (weights) of the machine--learning powered Maximally
Predictable Portfolio (MPP) are similar to weighting by the Kelly criterion.
Byrnes and Barnett~\cite{byrnesBarnett} went further to generalize the Kelly
Criterion with a continuous probability distribution modelling the price
movement of the stock, while Lototsky and Pollok~\cite{lototskyPollok}
rigorously analyzed the high--frequency limit of the model, concluding that a
high--frequency strategy can lead to a more aggressive betting strategy.
Proskurnikov and Barmish~\cite{proskurnikovBarmish} investigated the possibility
of having a non--linear control for the Kelly betting, assuming a bounded set
for the input probabilities instead of assuming that they are known in advance.
The Kelly Criterion was recently even extended to the quantum realm
by Meister and Price~\cite{meisterPrice}, where in addition to the uncertainty
related to flipping a coin, there is an uncertainty in measurement. Perhaps the
most related reference to this work is the paper by Jim\'{e}nez et
al.~\cite{jimenezEtAl} where they extended the Kelly Criterion to both
multivariate and simultaneous bets and applied it to sports betting in the
English Premier League. This work largely follows the approach by Jim\'{e}nez et
al.~\cite{jimenezEtAl}, but applied to a fundamentals based, focused investing:
\begin{enumerate}
    \item Multiple companies (stocks) are considered simultaneously,
    \item Each with an arbitrary number of scenarios,
    \item Multiple constraints for modelling our preference towards no shorting,
    no use of leverage, and providing a maximum value for the risk of permanent
    loss of capital,
    \item A fundamentals--based analysis with a very long time horizon. After
    all, a stock is an ownership share of business~\cite{intelligentInvestor},
    in which time--scales are significantly larger than the time--scales of
    day--to--day stock price fluctuations.
\end{enumerate}

@@

During the mathematical derivation presented in \autoref{sec:mathematics}, an
assumption is made that the number of bets is very high (tends to infinity).
This work does not try to justify this assumption in a strong mathematical
form~\footnote{It is on the authors' TODO list to try and prove this.}. Here's a
soft, non-mathematical reasoning on why the authors think this is fine: The
assumption is made in order to write the equations in terms of probabilities
instead of the number of outcomes divided by the number of bets. Therefore, as
long as the input probabilities are \textit{conservatively} estimated, the
framework should still be valid. This essentially represents the most important
margin of safety~\cite{intelligentInvestor}.

@@

    \item Maximum allowable permanent capital loss constraint ensures that the
    worst--case outcome does not exceed losing a specified amount of capital
    with a specified probability. Adding this constraint means: "Under a
    worst--case scenario, I am comfortable permanently losing $K$ (e.g. 25\%) of
    my capital with probability $P$ (e.g. 0.1\%).":
        \begin{equation}
        \label{eq:14}
            \sum_j^{N_c} f_j \min(p_{ij} k_{ij}) \ge P \cdot K \rightarrow
            -\sum_j^{N_c} \left(f_j \min(p_{ij} k_{ij}) \right) + P \cdot K + s = 0
        \end{equation}
    \noindent where $\min(p_{ij} k_{ij})$ is the worst--case outcome across all
    scenarios for the $j$-th candidate company. Here, the minimum returns
    $k_{ij}$ and the maximum worst--case return $K$ are both negative by
    convention, indicating a loss, hence the $\ge$ sign.
\end{enumerate}

@@

@@

\noindent The observation about excessive diversification is inline with the
thoughts from the Poor Charlie's Almanack~\cite{almanack} and one of the
lectures from Li Lu that the authors frequently watch~\cite{liLuLecture}. In
addition, it seems that recent research is also drawing similar conclusions
about excessive diversification~\cite{jamesMenzies,arvanitisEtAl}, although
using different investing frameworks. The study by James and
Menzies~\cite{jamesMenzies} is particularly interesting to the authors since the
data suggests that strong investors should hold concentrated portfolios, which
is in--line with practical observations of famous fundamentals--based investors:
Warren Buffett~\cite{berkshireLetters}, Mohnish Pabrai, Nick
Sleep~\cite{nomadLetters}, Pulak Prasad~\cite{darwinInvesting}, and others.
